Trending assessment for nowcasts, measurements, and forecasts

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \parencite{Cohen1960} or those listed in \textcite[Table 3.3]{Jolliffe2012}.
Cohen's $\kappa$ is usually used to measure inter-rater agreement and takes into account the ratio of occurred agreement and the probability of agreement by chance.
Cohen's $\kappa$ reduces to rescaling the trending ratio in the case of a $2\times2$ table and balanced outcomes.
Balanced outcome refers to approximately equal occurrences of values smaller and greater than zero for both $\diffx$ and $\diffy$.
Unbalanced outcomes of the differences are unlikely in our setting as they are obtained from differencing time series data.
Nevertheless, if the number of positive and negative $\diffy$ differs widely, the use of unbalanced-data-aware measures should be considered.

All measures can also be evaluated as a rolling estimate to detect changes in performance over time.
A rolling estimate is a sequence of estimates, each computed on a length-$w$-window of the data.
For the trending ratio, a rolling estimate with a backward-looking window is given by
\begin{equation*}
    \acc_t (\diffx, \diffy) \coloneqq \frac{\sum_{t^* = t-w + 1}^{t} \ind{\diffxt[t^*] \diffyt[t^*] > 0}}{w} \quad t = w-1, \dots, T.\label{eq:acc_rolling}
\end{equation*}
Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
Thus, the yearly course of the prediction's trending ratio can be detected.
The trending ability of the signal has a strong yearly, sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
This seasonal behavior cannot be observed in the two four-quadrant plots as the high number of points overlays it or in the trending ratio.

@@

\todo[inline]{Bsp. für einfache Exclusion area mit rectangle und resulting measures; Bsp. für andere shapes und wie dann $e$ gestaltet wäre}

>>>>

The use of those estimators could encourage the predictors to exploit imbalances in the number of positive and negative changes. 
A large difference in $\sum_{\mathcal} \ind{\diffyt > 0}$ and $\sum_{\mathcal} \ind{\diffyt > 0}$ is unlikely in the trending setting as $\diffy$ is obtained from differencing time series data.
Nevertheless, if the number of positive and negative $\diffy$ differs widely, the use of unbalanced-data-aware measures should be considered.
There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \parencite{Cohen1960} or those listed in \textcite[Table 3.3]{Jolliffe2012}.
Cohen's $\kappa$ is usually used to measure inter-rater agreement and takes into account the ratio of occurred agreement and the probability of agreement by chance.
Cohen's $\kappa$ reduces to rescaling the trending ratio in the case of a $2\times2$ table and balanced outcomes.

All measures can also be evaluated as a rolling estimate to detect changes in performance over time.
A rolling estimate is a sequence of estimates, each computed on a length-$w$-window of the data.
For the trending ratio, a rolling estimate with a backward-looking window is given by
\begin{equation*}
    \acc_t (\diffx, \diffy) \coloneqq \frac{\sum_{t^\star = t-w + 1}^{t} \ind{\diffxt[t^\star] \diffyt[t^\star] > 0}}{w} \quad t = w-1, \dots, T.\label{eq:acc_rolling}
\end{equation*}
Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
Thus, the yearly course of the prediction's trending ratio can be detected.
The trending ability of the signal has a strong yearly, sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
This seasonal behavior cannot be observed in the two four-quadrant plots as the high number of points overlays it or in the trending ratio.

@@
